---
title: "Modularity Calculation"
output:
  html_notebook:
    code_folding: null
  word_document: default
  html_document:
    df_print: paged
date: "2024-04-01"
---
#### Load libraries
```{r eval=FALSE, warning=FALSE}
library(R.matlab)
library(readxl)
library(igraph)
library(ggraph)
library(writexl)
```

### <span style="color: blue;">Set up enviorment</span>

#### Clear Environment and Console
```{r}
rm(list = ls()) # Clears environment
cat("\f")       # Clears console
```

#### Set working directory
```{r setup}
knitr::opts_knit$set(root.dir = normalizePath("/Users/astridrr/Desktop/Courses/Independent_study/Code")) # This sets the working directory to your specified path.
```

### <span style="color: blue;">Prepare Data</span>

#### Import data
Connectivity matrices and ROI names in .mat file
```{r}
data <- readMat("Schaefer_400ROI_Data.mat") # Load connectivity matrices and ROI names from a .mat file
ROI_names <- data[["ROI.names"]] # Extract ROI names from the loaded data
stacked_Z <- data[["stacked.Z"]] # Extract the stacked connectivity matrices for all subjects
```
ROI details on .xlsx file
```{r}
nodes <- read_excel("Schaefer_17Network_400ROIs_detail.xlsx") # Load ROI details from an Excel file
nodes <- nodes[,c("Order","Label","Network_Name")] # Retain only the columns "Order", "Label", and "Network_Name" for analysis
```

#### Process data
Extract subject data from stacked matrix
```{r}
num_subs <- dim(stacked_Z)[3] # Determine the number of subjects based on the third dimension of the stacked matrix
all_subs_raw <- list() # Initialize a list to hold the matrices for each subject

# Stores each subject's matrix on a list
for (n in 1:num_subs){
    sub <- stacked_Z[,,n] # Extract the subject's matrix from the staked matrix
    all_subs_raw[[n]] <- sub # Stores the subject's matrix on a list
}
```
Eliminate NaNs and negative values from the matrices
```{r}
all_subs <- list() # Initialize a list for storing cleaned subject matrices

for (n in 1:num_subs){
    sub <- all_subs_raw[[n]]
    sub[is.nan(sub)] <- 0 # Replace NaN values with 0 in the matrix
    sub[sub < 0] <- 0 # Set negative values to 0 assuming only non-negative correlations are of interest.
    all_subs[[n]] <- sub # Stores the cleaned subject's matrix on a list
}
```

#### Convert to `igraph`
```{r}
subs_graphs <- list() # Initialize empty list to store subject's igraphs

for (n in 1:num_subs){
    sub <- all_subs[[n]]
    sub <- graph_from_adjacency_matrix(sub, mode = "max", weighted = TRUE)
    subs_graphs[[n]] <- sub
}
```

### <span style="color: blue;">Generating random networks</span>
Generate a random network preserving the degree distribution of the original networks.
```{r}
#rand_graphs <- list()
#
#for (n in 1:num_subs){
#    sub <- subs_graphs[[n]]
#    deg <- degree(sub)
#    rand_graphs_sub <- list()
#    for (m in 1:100) {
        # Generate a random graph with the same degree sequence
#        rand <- sample_degseq(deg, method = "simple.no.multiple")
        # Add the random graph to the list
#        rand_graphs_sub[[m]] <- rand
#    }
    # Calculate the mean of the random graphs for this subject
#    mean_rand_graph <- Reduce(`+`, rand_graphs_sub) / length(rand_graphs_sub)
    # Add the mean random graph to the overall list
#    rand_graphs[[n]] <- mean_rand_graph
#}
```

### <span style="color: blue;">Detect communities and calculate modularity</span>

#### Detect communities with Louvain method
```{r}
sub_communities <- list()

for (n in 1:num_subs){
    sub <- subs_graphs[[n]]
    sub <- cluster_louvain(sub, weights = E(sub)$weight)
    sub_communities[[n]] <- sub
}
```

#### Calculate modularity
```{r}
mod_lv <- list()

for (n in 1:num_subs){
    sub_g <- subs_graphs[[n]]
    sub_c <- sub_communities[[n]]
    sub <- modularity(sub_g, sub_c$membership)
    mod_lv[[n]] <- sub
}
```

#### Save modularity values
```{r}
sub_mod_lv <- data.frame(mod_lv = unlist(mod_lv))
sub_mod_lv$ID <- 1:252
write_xlsx(sub_mod_lv, "/Users/astridrr/Desktop/Courses/Independent_study/Code/sub_mod_lv.xlsx")
```

#### Visualize modularity
```{r}
#plot(communities, net, vertex.size = 5) # Visualize the detected communities within the network.
#plot(subgraph(net, sample(V(net), 50))) # Plot a sub sample
```

### <span style="color: blue;">Modularity based on Schaefer 17 networks</span>

#### Calculate modularity
```{r}
# Extract the network labels for each node
subnet <- nodes$Label

mod_sch <- list()

for (n in 1:num_subs){
    sub_g <- subs_graphs[[n]]
    sub <- modularity(sub_g, membership = subnet, weights = E(sub)$weight)
    mod_sch[[n]] <- sub
}
# Calculate modularity with predefined communities
# modularity_val <- modularity(net, membership = subnet)
# print(paste("Modularity (Schaefer 17 Networks) = ", modularity_val))
```

#### Visualization of Network Communities
```{r}
#ggraph(net) + geom_edge_link(aes(alpha=.5)) + geom_node_point(aes(color=factor(subnet))) + theme_graph()
```






