---
title: "Modularity Statistics"
output:
  html_notebook:
    code_folding: null
  word_document: default
  html_document:
    df_print: paged
date: "2024-04-01"
---
#### Load libraries
```{r eval=FALSE, warning=FALSE}
library(readxl)
library(tidyverse)
library(magrittr)
```

### <span style="color: blue;">Set up enviorment</span>

#### Clear Environment and Console
```{r}
rm(list = ls()) # Clears environment
cat("\f")       # Clears console
```

#### Set working directory
```{r setup}
knitr::opts_knit$set(root.dir = normalizePath("/Users/astridrr/Desktop/Courses/Independent_study/Code")) # This sets the working directory to your specified path.
```

### <span style="color: blue;">Prepare Data</span>

#### Import data
```{r}
dat <- read_excel("merge_data.xlsx")
mod_lv <- read_excel("sub_mod_lv.xlsx")
```

#### Merge data 
```{r}
names(mod_lv) <- c("mod_lv","Matrix_Num") # Changes the variable names to match dat

data <- merge(dat, mod_lv, by = "Matrix_Num")
```

#### Clean data
```{r}
data <- data[data$Dx!=4 & data$Dx!=5,] # Excludes subjects under categories 4 and 5
```

#### Recode variables
```{r}
data$Sex <- car::recode(data$Sex,"'Female'=0;'Male'=1") # Recodes Females and Males as random variables 0 and 1
data$SMOKSTA <- car::recode(data$SMOKSTA, "'2'=1")
data$Amyloid <- car::recode(data$Amyloid, "'Negative'=0;'Positive'=1;'Inconclusive'=9;NA=9")
```

#### Data subset
```{r}
odata <-data[data$AgeatMRI>53,] # Extracts subjects ≥54 years old from the data
adata <-data[data$Amyloid!=9,]
```

### <span style="color: blue;">Exploratory data analysis</span>

#### Modularity through age

Previous studies have shown that modularity decreases with age.
https://doi.org/10.1089/brain.2014.0286


```{r}
mdl <- "mod_lv ~ 1 + AgeatMRI"
lm <- lm(mdl, data)
lmAov <- lm %>% car::Anova(.,type=2) %T>% print()
lmSum <- lm %>% summary() %>% {.$coefficients} %>% cbind(confint(lm)[row.names(.),]) %>% print()
cat("Multiple R-squared:", summary(lm)$r.squared,"\nAdjusted R-squared:", summary(lm)$adj.r.squared, "\n")
```
Visualizing the data
```{r}
fig <- ggplot(data, aes(x = AgeatMRI, y = mod_lv)) +
    geom_point(aes(col = factor(Dx), alpha = 0.8)) + theme_bw() + 
    geom_smooth(method = "lm", se = FALSE, color = "black") +
    labs(x = "Age", y = "Modularity", colour = "Legend") +  # Set x and y axis labels
    scale_colour_hue(labels = c("1" = "CN", "2" = "MCI", "3" = "AD"))
    
    print(fig)
```

#### Modularity in diagnosis groups

```{r}
mdl <- "mod_lv ~ 1 + AgeatMRI + Dx"
lm <- lm(mdl, data)
lmAov <- lm %>% car::Anova(.,type=2) %T>% print()
lmSum <- lm %>% summary() %>% {.$coefficients} %>% cbind(confint(lm)[row.names(.),]) %>% print()
cat("Multiple R-squared:", summary(lm)$r.squared,"\nAdjusted R-squared:", summary(lm)$adj.r.squared, "\n")
```
Visualizing the data
```{r}
fig <- ggplot(data, aes(x = factor(Dx), y = mod_lv, fill= factor(Dx))) +
    geom_violin() + theme_bw() + 
    stat_summary(fun.data = "mean_cl_boot", geom = "pointrange") + # Add dot representing mean with sd
    labs(x = "Diagnosis", y = "Modularity") +  # Set x and y axis labels
    scale_fill_discrete(guide = FALSE) +  # Remove legend
    scale_x_discrete(labels = c("1"= "HC", "2" = "MCI", "3" = "AD"))  # Set custom tick marks

print(fig)
```

#### Modularity in smoking status

```{r}
mdl <- "mod_lv ~ 1 + AgeatMRI + SMOKSTA"
lm <- lm(mdl, data)
lmAov <- lm %>% car::Anova(.,type=2) %T>% print()
lmSum <- lm %>% summary() %>% {.$coefficients} %>% cbind(confint(lm)[row.names(.),]) %>% print()
cat("Multiple R-squared:", summary(lm)$r.squared,"\nAdjusted R-squared:", summary(lm)$adj.r.squared, "\n")
```
Visualizing the data
```{r}
fig <- ggplot(data, aes(x = factor(SMOKSTA), y = mod_lv, fill= factor(SMOKSTA))) +
    geom_violin() + theme_bw() + 
    stat_summary(fun.data = "mean_cl_boot", geom = "pointrange") + # Add dot representing mean with sd
    labs(x = "Smoking status", y = "Modularity") +  # Set x and y axis labels
    scale_fill_discrete(guide = FALSE) +  # Remove legend
    scale_x_discrete(labels = c("0"= "Non-Smoker", "1" = "Smoker"))  # Set custom tick marks

print(fig)
```

#### Modularity in amyloid status 

```{r}
mdl <- "mod_lv ~ 1 + AgeatMRI + Amyloid"
lm <- lm(mdl, adata)
lmAov <- lm %>% car::Anova(.,type=2) %T>% print()
lmSum <- lm %>% summary() %>% {.$coefficients} %>% cbind(confint(lm)[row.names(.),]) %>% print()
cat("Multiple R-squared:", summary(lm)$r.squared,"\nAdjusted R-squared:", summary(lm)$adj.r.squared, "\n")
```
Visualizing the data
```{r}
fig <- ggplot(adata, aes(x = factor(Amyloid), y = mod_lv, fill= factor(Amyloid))) +
    geom_violin() + theme_bw() + 
    stat_summary(fun.data = "mean_cl_boot", geom = "pointrange") + # Add dot representing mean with sd
    labs(x = "Amyloid status", y = "Modularity") +  # Set x and y axis labels
    scale_fill_discrete(guide = FALSE) +  # Remove legend
    scale_x_discrete(labels = c("0"= "Negative", "1" = "Positive"))  # Set custom tick marks

print(fig)
```




